---
title: "ANOVA Report"
output:
  html_document:
    theme: united
    toc: true
    toc_float: true
params:
  model_formula: NULL
  dataset: NULL
  anova_output: NULL
  tukey_plot: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("helpers.R", local = TRUE)
library(ggplot2)
library(lsr)
library(htmltools)
library(dplyr)
library(agricolae)
library(htmlTable)
library(car)
library(nortest)
library(gghalves)
```

##  Model Formula

**`r as.character(params$model_formula)`**

## 📄 Dataset Preview

```{r dataset-preview}
knitr::kable(params$dataset, caption = "Full Dataset")
```

##  ANOVA Table

```{r anova-table, results='asis'}
params$anova_output$htmlTable
```

##  Summary

```{r summary-card, results='asis'}
depvar <- all.vars(params$model_formula)[1]
indvar <- all.vars(params$model_formula)[-1]
HTML(do.call(summary_card, list(model = params$anova_output, df = params$dataset, depvar = depvar, indvar = indvar)))
```

##  Histogram of Dependent Variable

```{r histogram}
dep <- all.vars(params$model_formula)[1]
ggplot(params$dataset, aes_string(x = dep)) +
  geom_histogram(fill = "lightgreen", color = "black", bins = 30) +
  theme_minimal() +
  labs(title = "Histogram of Dependent Variable", x = dep, y = "Count")
```

##  Boxplot of Dependent Variable

```{r boxplot}
ind <- paste(all.vars(params$model_formula)[-1], collapse = ":")
ggplot(params$dataset, aes_string(x = ind, y = all.vars(params$model_formula)[1])) +
  geom_boxplot(fill = "#a0c4ff") +
  theme_minimal() +
  labs(title = "Boxplot of Dependent Variable by Group", x = "Group", y = all.vars(params$model_formula)[1])
```

##  QQ Plot of Residuals

```{r qq-plot}
model <- lm(params$model_formula, data = params$dataset)
qqnorm(model$residuals)
qqline(model$residuals, col = "red")
```

##  Residuals vs Fitted

```{r residuals-vs-fitted}
plot(model$fitted.values, model$residuals,
     xlab = "Fitted Values", ylab = "Residuals",
     main = "Residuals vs Fitted")
abline(h = 0, col = "red", lty = 2)
```


##  Multi-vari Chart

```{r multivari}
if (length(all.vars(params$model_formula)) == 3) {
  dep <- all.vars(params$model_formula)[1]
  ind1 <- all.vars(params$model_formula)[2]
  ind2 <- all.vars(params$model_formula)[3]
  df <- params$dataset
  df[[ind1]] <- factor(df[[ind1]])
  df[[ind2]] <- factor(df[[ind2]])
  ggplot(df, aes_string(x = ind1, y = dep, color = ind2, group = ind2)) +
    stat_summary(fun = mean, geom = "point", size = 3) +
    stat_summary(fun = mean, geom = "line") +
    theme_minimal() +
    labs(title = "Multi-vari Chart", x = ind1, y = dep, color = ind2)
}
```


##  Violin Plot

```{r violin}
dep <- all.vars(params$model_formula)[1]
ind <- paste(all.vars(params$model_formula)[-1], collapse = ":")
df <- params$dataset
df$Group <- interaction(df[, all.vars(params$model_formula)[-1]], drop = TRUE)
ggplot(df, aes(x = Group, y = .data[[dep]], fill = Group)) +
  geom_violin(alpha = 0.5) +
  geom_jitter(width = 0.15, alpha = 0.5, color = "gray30") +
  theme_minimal() +
  labs(title = "Violin Plot", x = "Group", y = dep)
```

##  Raincloud Plot

```{r raincloud}
df <- params$dataset
df$Group <- interaction(df[, all.vars(params$model_formula)[-1]], drop = TRUE)
ggplot(df, aes(x = Group, y = .data[[dep]], fill = Group)) +
  gghalves::geom_half_violin(side = "l", alpha = 0.5) +
  geom_boxplot(width = 0.1, outlier.shape = NA) +
  gghalves::geom_half_point(side = "r", alpha = 0.4) +
  theme_minimal() +
  labs(title = "Raincloud Plot", x = "Group", y = dep)
```
##  Effect Sizes (η²)

```{r eta-squared}
eta_vals <- etaSquared(aov(model), type = 3)
eta_df <- as.data.frame(eta_vals)
eta_df$Term <- rownames(eta_df)
ggplot(eta_df, aes(x = reorder(Term, eta.sq), y = eta.sq)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  theme_minimal() +
  labs(title = "Effect Sizes (η²)", x = "Effect", y = "η²")
```


##  Shapiro-Wilk & Lilliefors

```{r shapiro-lillie}
resids <- residuals(model)
hist(resids, breaks = 20, main = "Histogram of Residuals (Shapiro-Wilk)",
     xlab = "Residuals", col = "lightblue", border = "white")
mtext(sprintf("Shapiro-Wilk p = %.3f", shapiro.test(resids)$p.value), side = 3, line = 0.5)

hist(resids, breaks = 20, probability = TRUE,
     main = "Lilliefors Test: Residuals vs Normal", col = "lightpink", border = "white",
     xlab = "Residuals")
curve(dnorm(x, mean(resids), sd(resids)), col = "darkred", lwd = 2, add = TRUE)
mtext(sprintf("Lilliefors p = %.3f", nortest::lillie.test(resids)$p.value), side = 3, line = 0.5)
```

##  Normality Tests by Group

```{r by-group-tests}
group <- interaction(params$dataset[, all.vars(params$model_formula)[-1]], drop = TRUE)
dep <- all.vars(params$model_formula)[1]
df <- data.frame(Group = group, Value = params$dataset[[dep]])

shapiro_vals <- by(df$Value, df$Group, function(x) {
  if (length(x) >= 3 && length(unique(x)) > 2) {
    tryCatch(shapiro.test(x)$p.value, error = function(e) NA)
  } else NA
})
df_shapiro <- data.frame(Group = names(shapiro_vals), P_value = as.numeric(shapiro_vals))
ggplot(df_shapiro, aes(x = Group, y = P_value)) +
  geom_bar(stat = "identity", fill = "#5dade2") +
  geom_hline(yintercept = 0.05, linetype = "dashed", color = "red") +
  theme_minimal() +
  labs(title = "Shapiro-Wilk Test by Group", y = "p-value", x = "Group")

lillie_vals <- by(df$Value, df$Group, function(x) {
  if (length(x) >= 4 && length(unique(x)) > 2) {
    tryCatch(nortest::lillie.test(x)$p.value, error = function(e) NA)
  } else NA
})
df_lillie <- data.frame(Group = names(lillie_vals), P_value = as.numeric(lillie_vals))
ggplot(df_lillie, aes(x = Group, y = P_value)) +
  geom_bar(stat = "identity", fill = "#48c9b0") +
  geom_hline(yintercept = 0.05, linetype = "dashed", color = "red") +
  theme_minimal() +
  labs(title = "Lilliefors Test by Group", y = "p-value", x = "Group")
```
## 🔹 Tukey HSD Plot
  
```{r tukey-hsd-plot, echo=FALSE, fig.align='center', fig.cap='Tukey HSD Confidence Intervals'}
# Extract variable names
vars <- all.vars(params$model_formula)
depvar <- vars[1]
indvar <- vars[2]

# Work on a local copy to avoid modifying params
local_df <- params$dataset
local_df[[indvar]] <- factor(trimws(as.character(local_df[[indvar]])))

# Fit the model
aov_model <- aov(params$model_formula, data = local_df)

# Run and plot Tukey HSD
tukey_result <- TukeyHSD(aov_model)
plot(tukey_result, las = 1, col = "steelblue")

```

------------------------------------------------------------------------
##  Group Means with 95% CI & Tukey Grouping
  
```{r group-means-ci, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(agricolae)

# Extract variables
vars <- all.vars(params$model_formula)
dep <- vars[1]
ind <- vars[2]
df <- params$dataset
df[[ind]] <- factor(trimws(tolower(as.character(df[[ind]]))))

# Prepare summary statistics
summary_df <- df %>%
  group_by(.data[[ind]]) %>%
  summarise(
    Mean = mean(.data[[dep]], na.rm = TRUE),
    N = n(),
    SD = sd(.data[[dep]], na.rm = TRUE),
    SEM = SD / sqrt(N),
    .groups = 'drop'
  ) %>%
  mutate(
    CI_lower = Mean - 1.96 * SEM,
    CI_upper = Mean + 1.96 * SEM
  )

# Tukey grouping
aov_model <- aov(params$model_formula, data = df)
tukey_result <- agricolae::HSD.test(aov_model, trt = ind, group = TRUE)
groups <- as.data.frame(tukey_result$groups)
groups[[ind]] <- rownames(groups)
summary_df <- merge(summary_df, groups[, c(ind, "groups")], by = ind)

# Plot
ggplot(summary_df, aes_string(x = ind, y = "Mean")) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper), width = 0.2) +
  geom_text(aes(label = groups, y = CI_upper + 0.1 * max(CI_upper)), vjust = 0) +
  theme_minimal() +
  labs(title = "Group Means ± 95% CI with Tukey Grouping", x = ind, y = dep)
```

------------------------------------------------------------------------
  
## Homogeneity Tests (Levene & Bartlett)
  
```{r lev-bart-table, results='asis'}
my_check_homogeneity_of_column(params$dataset, all.vars(params$model_formula)[1], all.vars(params$model_formula)[2])
```

------------------------------------------------------------------------
  
##  Standard Deviation per Group
  
```{r sd-plot}
library(dplyr)
library(ggplot2)
df <- params$dataset
dep <- all.vars(params$model_formula)[1]
ind <- all.vars(params$model_formula)[2]

sd_df <- df %>%
  group_by(.data[[ind]]) %>%
  summarise(SD = sd(.data[[dep]], na.rm = TRUE)) %>%
  ungroup()

ggplot(sd_df, aes_string(x = ind, y = "SD")) +
  geom_bar(stat = "identity", fill = "darkorange") +
  theme_minimal() +
  labs(title = "Standard Deviation per Group", x = ind, y = "Standard Deviation")
```

------------------------------------------------------------------------
  
##  Residual Spread by Group
  
```{r residual-spread-plot}
model <- lm(params$model_formula, data = params$dataset)
plot_data <- data.frame(
  Group = params$dataset[[all.vars(params$model_formula)[2]]],
  AbsResidual = abs(residuals(model))
)

ggplot(plot_data, aes(x = Group, y = AbsResidual)) +
  geom_boxplot(fill = "#fbc531") +
  theme_minimal() +
  labs(
    title = "Spread of Absolute Residuals by Group (Levene Visual)",
    x = "Group", y = "Absolute Residual"
  )
```